Simple repression draft

### 2020-06-01 doc/section_03_beyond_means.tex
MRM pass through entire main text. Removed side notes. Still lots left

\subsection{A one-state promoter with bursting is the ``best'' model of
constitutive promoters}

@@

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/main/fig02.pdf}
\caption{\textbf{Comparison of different models for noise in the constitutive promoter.}
(A) The left column depicts various plausible models for the dynamics of
constitutive promoters. In model (1), transcripts are produced in a Poisson
process~\cite{Sanchez2013, Jones2014}. Model (2) features explicit modeling of
RNAP binding/unbinding kinetics~\cite{Phillips2015a}. Model (3) is a more
detailed generalization of model (2), treating transcription initiation as a
multi-step process proceeding through closed and open
complexes~\cite{Mitarai2015}. Model (4) is somewhat analogous to (2) except with
the precise nature of active and inactive states left
ambiguous~\cite{Peccoud1995, Shahrezaei2008, Razo-Mejia2020}. Finally, model (5)
can be viewed as a certain limit of model (4) in which transcripts are produced
in bursts, and initiation of bursts is a Poisson process. 
% \mmnote{Haven't found model 5 formulated quite like this in the literature,
% usually burstiness is handled like model 4, but I should probably search a bit
% more.} 
The right column shows the Fano factor $\nu$ (variance/mean) for each model.
Note especially the crucial diagnostic: (2) and (3) have $\nu$ strictly below 1,
while only for (4) and (5) can $\nu$ exceed 1. Models with Fano factors $\le 1$
cannot produce the FISH data observed in part (B) without introducing
additional assumptions and model complexity. (B) Data from \cite{Jones2014}.
Mean mRNA count vs. Fano factor (variance/mean) for different promoters as
determined with single-molecule mRNA FISH. The colorbar indicates the predicted
binding affinity of RNAP to the promoter sequence as determined in
\cite{Brewster2012}. Numbers serve for cross comparison with data presented in 
Figure 3.}
\label{fig2:constit_cartoons}
\end{figure}

@@

The simplest model of constitutive expression that we can imagine is shown as
model 1 in Figure~\ref{fig2:constit_cartoons} and assumes that transcripts are
produced as a Poisson process from a single promoter state. This is the picture
from Jones et.\ al.~\cite{Jones2014} that was used to interpret a systematic
study of gene expression noise over a series of promoters designed to have
different strengths. This model insists that the ``true'' mRNA distribution is
Poisson, implying the Fano factor $\nu$ must be 1. In~\cite{Jones2014}, the
authors carefully attribute measured deviations from Fano = 1 to intensity
variability in fluorescent spots, gene copy number variation, and copy number
fluctuations of the transcription machinery, e.g., RNAP itself. In this picture,
the master equation makes no appearance, and all the corrections to Poisson
behavior are derived as additive corrections to the Fano factor. For disproving
the ``universal noise curve'' from So et al~\cite{So2011}, this picture was
excellent. It is appealing in its simplicity, with only two parameters, the
initiation rate $r$ and degradation rate $\gamma$. Since $\gamma$ is
independently known from other experiments, and the mean mRNA copy number is
$r/\gamma$, $r$ is easily inferred from data. In other words, the model is not
excessively complex for the data at hand. But for many interesting questions,
for instance in the recent work~\cite{Razo-Mejia2020}, knowledge of means and
variances alone is insufficient, and a description of the full distribution of
molecular counts is necessary. For this we need a (slightly) more complex model
than model 1 that would allow us to incorporate the non-Poissonian features of
constitutive promoters directly into a master equation formulation.

@@

Our second model of constitutive transcription posits an architecture in which
the promoter is either empty or bound by RNAP ~\cite{Phillips2015a,
Phillips2019}. Here,  as shown in model 2 of Figure~\ref{fig2:constit_cartoons},
transcription initiation results in a state transition from the bound to the
unbound state, reflecting the microscopic reality that an RNAP that has begun to
elongate a transcript is no longer available at the start site to begin another.
As shown in Appendix~\ref{sec:non_bursty}, the Fano factor in this model is
given by
\begin{align}
    \nu = 1 -
        \frac{r\rate{k}{P}{+}}
            {\left(\rate{k}{P}{+} + \rate{k}{P}{-} + r\right)
             \left(\gamma + \rate{k}{P}{+} + \rate{k}{P}{-} + r\right)}.
\end{align}
The problem with this picture is that the Fano factor is always $<1$. To make
 contact with the experimental reality of $\nu>1$, clearly some corrections will
 be needed. While this model adds an appealing element of microscopic reality,
 we are forced to reject it as the additional complexity is unable to capture
 the phenomenology of interest. Obviously the promoter state does in fact
 proceed through cycles of RNAP binding, initiating, and elongating, but it
 seems that the super-Poissonian noise in mRNA copy number we want to model must
 be governed by other features of the system.

@@

\marginpar{\it RP: very fun next paragraph - I like the argument, need to think more}
In fact, we suspect \textit{any} model in which transcription proceeds through a
multistep cycle must necessarily have $\nu<1$. The intuitive argument compares
the waiting time distribution to traverse the cycle with the waiting time for a
Poisson promoter (model 1) with the same mean time. The latter is simply an
exponential distribution. The former is a convolution of multiple exponentials,
and intuitively should be more peaked with a smaller fractional width than an
exponential with the same mean.
\footnote{This can be made more precise with a result from~\cite{Stewart2007},
who showed that the convolution of multiple gamma distributions (of which the
exponential distribution is a special case) is, to a very good approximation,
also gamma distributed. Using their Eq.~2 for the distribution of the
convolution, with shape parameters set to 1 to give exponential distributions,
the total waiting time distribution has a ratio of variance to squared mean
$\sigma^2/\mu^2 = \sum_i k_i^2/\left(\sum_i k_i\right)^2$, where the $k_i$ are
the rates of the individual steps. Clearly this is less than 1 and therefore the
total waiting time distribution is narrower than the corresponding exponential.}
A less disperse waiting time distribution means transcription initations are
more uniformly distributed in time relative to a Poisson process. Hence the
distribution of mRNA over a population of cells should be less variable compared
to Poisson, giving $\nu<1$.
\footnote{This last step, while intuitive, can be argued by analogy to photon
statistics where antibunching gives rise to sub-Poissonian noise~\cite{Paul1982,
Zou1990}. Although loopholes exist, we do not expect they apply for our problem.
Nevertheless we refrain from elevating this cycle/sub-Poissonian equivalence to
a ``theorem.'' \mmnote{Antibunching is the ``obvious'' analogy for me but it'd
be totally out of left field for most readers, any suggestions on a more
comprehensible reference??}}

Regardless of the merits of this model in describing the noise properties of
constitutive transcription initiation, it ultimately fails the simplest
quantitative feature of the data, namely that the Fano factor $> 1$ and hence
we must discard this mechanistic picture and search elsewhere.

@@

While this seems promising, there is a major drawback for our purposes which was
already uncovered by the authors of~\cite{Razo-Mejia2020}: the statistical
inference problem is nonidentifiable, in the sense described in Section 4.3
of~\cite{Gelman2013}. What this means is that it is impossible to infer the
parameters $r$ and $k^-$ from the FISH data of~\cite{Jones2014} (as shown in
Fig.~S2 of~\cite{Razo-Mejia2020}). Rather, only the ratio $r/k^-$ could be
inferred. In that work, the problem was worked around with an informative prior
on the ratio $k^-/k^+$. That approach is unlikely to work here, as, recall, our
entire goal in modeling constitutive expression is to use it as the basis for a
yet more complicated model, when we add on repression. But adding more
complexity to a model that is already poorly identified is a fool's errand, so
we will explore one more potential model.

@@

The final model we consider is 
inspired by the failure mode of model 4. The key observation above
was that, as found in~\cite{Razo-Mejia2020}, only two parameters, $k^+$ and the
ratio $r/k^-$, could be directly inferred from the FISH data
from~\cite{Jones2014}. So let us take this seriously and imagine a model where
these are the only two model parameters. What would this model look like?

@@

Let us now verify this intuition with a precise derivation to check that $r/k^-$
is in fact the mean burst size and to obtain the full burst size distribution.
Consider first a constant, known dwell time $T$ in the active state. Transcripts
are produced at a rate $r$ per unit time, so the number of transcripts $n$
produced during $T$ fits exactly the ``story'' for the Poisson distribution,
i.e.,
\begin{equation}
    P(n\mid T) = \frac{(rT)^n}{n!} \exp(-rT).
\end{equation}
Since the dwell time $T$ is unobservable, we actually want $P(n)$, the dwell
time distribution with no conditioning on $T$. Basic rules of probability theory
tell us we can write $P(n)$ in terms of $P(n\mid T)$ as
\begin{equation}
    P(n) =\int_0^\infty P(n\mid T) P(T) dT.
\end{equation}
But we know the dwell time distribution $P(T)$, which is exponentially
distributed according to
\begin{equation}
    P(T) = k^- \exp(-T k^-),
\end{equation}
so $P(n)$ can be written as
\begin{equation}
    P(n) = k^- \frac{r^n}{n!}
            \int_0^\infty T^n\exp[-(r + k^-)T]\,dT.
\end{equation}
A standard integral table shows $\int_0^\infty x^n e^{-ax}\,dx = n!/a^{n+1}$, so
\begin{equation}
    P(n) = \frac{k^- r^n}{(k^- + r)^{n+1}}
        = \frac{k^-}{k^- + r}
            \left(\frac{r}{k^- + r}\right)^n
        = \frac{k^-}{k^- + r}
            \left(1 - \frac{k^-}{k^- + r}\right)^n,
\end{equation}
which is exactly the geometric distribution with standard parameter
$\theta\equiv k^-/(k^- + r)$ and domain $n \in \{0, 1, 2, \dots\}$ (this is one
of two common conventions for the geometric distribution). The mean of the
geometric distribution, with this convention, is
\begin{align}
\langle n\rangle = \frac{1 - \theta}{\theta}
        = \left(1 - \frac{k^-}{k^- + r}\right)
                    \frac{k^- + r}{k^-}
        = \frac{r}{k^-},
\end{align}
exactly as we guessed intuitively above.

So in taking the limit $r,k^-\rightarrow\infty$, $r/k^-\equiv b$, we obtain a
model which effectively has only a single promoter state, which initiates bursts
at rate $k^+$ (transitions to the active state, in the model 4 picture). The
master equation for mRNA copy number $m$ as derived in
Appendix~\ref{sec:non_bursty} takes the form
\begin{align}
\begin{split}
{d \over dt}p(m,t) = & (m+1)\gamma p(m+1,t) - m\gamma p(m,t) \\
        &+ \sum_{m'=0}^{m-1} k_i p(m',t) \text{Geom}(m-m';b)
         - \sum_{m'=m+1}^\infty k_i p(m,t) \text{Geom}(m'-m;b),
\end{split}
\end{align}
where we use $k_i$ to denote the burst initiation rate, $\text{Geom}(n;b)$ is the
geometric distribution with mean~$b$, i.e., $\text{Geom}(n;b) =
\frac{1}{1+b}\left(\frac{b}{1+b}\right)^n$ (with domain over nonnegative
integers as above). The first two terms are the usual mRNA degradation terms.
The third term enumerates all ways the system can produce a burst of transcripts
and arrive at copy number $m$, given that it had copy number $m'$ before the
burst. The fourth term allows the system to start with copy number $m$, then
produce a burst and end with copy number $m'$. In fact this last sum has trivial
$m'$ dependence and simply enforces normalization of the geometric distribution.
Carrying it out we have
\begin{equation}
\begin{split}
{d \over dt} p(m,t) = & (m+1)\gamma p(m+1,t) - m\gamma p(m,t) \\
        &+ \sum_{m'=0}^{m-1} k_i p(m',t) \text{Geom}(m-m';b)
            - k_i p(m,t),
\end{split}
\end{equation}
We direct readers again to Appendix~\ref{sec:beyond_means} for further details.
This improves on model 4 in that now the parameters are easily inferred, as we
will see later, and have clean interpretations. The non-Poissonian features are
attributed to the empirically well-established phenomenological picture of
bursty transcription.

@@

Consider the time delay between the first and final RNAPs in a burst initiating
transcription (\textit{not} the time to complete transcripts, which potentially
could be much longer.) If this timescale is short compared to the typical search
timescale of transcription factors, then all is well. The estimates from
deHaseth et.\ al.~\cite{DeHaseth1998} put RNAP's diffusion-limited on rate
around $\sim\text{few}\times10^{-2}~\text{nM}^{-1}~\text{s}^{-1}$ and polymerase
loading as high as $1~\text{s}^{-1}$. Then for reasonable burst sizes of $<10$,
it is reasonable to guess that bursts might finish initiating on a timescale of
tens of seconds or less (with another 30-60 sec to finish elongation, but that
does not matter here). A transcription factor with typical copy number of order
10 (or less) would have a diffusion-limited association rate of order
$(10~\text{sec})^{-1}$ \mrm{cite Elf?}. Higher copy number TFs tend to have many
binding sites over the genome, which should serve to pull them out of
circulation and keep their effective association rates from rising too large.
Therefore, there is \textit{perhaps} a timescale separation possible between
transcription factor association rates and burst durations, but this assumption
could very well break down, so we will have to keep it in mind when we infer
repressor rates from the Jones et.\ al.\ FISH data later~\cite{Jones2014}.

>>>>

\subsection{Kinetic models for unregulated promoter noise}

@@

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/main/fig02.pdf}
\caption{\textbf{Comparison of different models for noise in the constitutive promoter.}
(A) The left column depicts various plausible models for the dynamics of
constitutive promoters. In model (1), transcripts are produced in a Poisson
process~\cite{Sanchez2013, Jones2014}. Model (2) features explicit modeling of
RNAP binding/unbinding kinetics~\cite{Phillips2015a}. Model (3) is a more
detailed generalization of model (2), treating transcription initiation as a
multi-step process proceeding through closed and open
complexes~\cite{Mitarai2015}. Model (4) is somewhat analogous to (2) except with
the precise nature of active and inactive states left
ambiguous~\cite{Peccoud1995, Shahrezaei2008, Razo-Mejia2020}. Finally, model (5)
can be viewed as a certain limit of model (4) in which transcripts are produced
in bursts, and initiation of bursts is a Poisson process. 
% \mmnote{Haven't found model 5 formulated quite like this in the literature,
% usually burstiness is handled like model 4, but I should probably search a bit
% more.} 
The right column shows the Fano factor $\nu$ (variance/mean) for each model.
Note especially the crucial diagnostic: (2) and (3) have $\nu$ strictly below 1,
while only for (4) and (5) can $\nu$ exceed 1. Models with Fano factors $\le 1$
cannot produce the single-cell data observed in part (B) without introducing
additional assumptions and model complexity. (B) Data from \cite{Jones2014}.
Mean mRNA count vs. Fano factor (variance/mean) for different promoters as
determined with single-molecule mRNA Fluorescence \textit{in situ}
Hybridization. The colorbar indicates the predicted binding affinity of RNAP to
the promoter sequence as determined in \cite{Brewster2012}. Numbers serve for
cross comparison with data presented in Figure 3.}
\label{fig2:constit_cartoons}
\end{figure}

@@

The simplest model of constitutive expression that we can imagine is shown as
model 1 in Figure~\ref{fig2:constit_cartoons} and assumes that transcripts are
produced as a Poisson process from a single promoter state. This is the picture
from Jones et.\ al.~\cite{Jones2014} that was used to interpret a systematic
study of gene expression noise over a series of promoters designed to have
different strengths. This model insists that the ``true'' steady-state mRNA
distribution is Poisson, implying the Fano factor $\nu$ must be 1.
In~\cite{Jones2014}, the authors carefully attribute measured deviations from
Fano = 1 to intensity variability in fluorescent spots, gene copy number
variation, and copy number fluctuations of the transcription machinery, e.g.,
RNAP itself. In this picture, the master equation makes no appearance, and all
the corrections to Poisson behavior are derived as additive corrections to the
Fano factor. For disproving the ``universal noise curve'' from So et
al~\cite{So2011}, this picture was excellent. It is appealing in its simplicity,
with only two parameters, the initiation rate $r$ and degradation rate $\gamma$.
Since $\gamma$ is independently known from other experiments, and the mean mRNA
copy number is $r/\gamma$, $r$ is easily inferred from data. In other words, the
model is not excessively complex for the data at hand. But for many interesting
questions, for instance in the recent work~\cite{Razo-Mejia2020}, knowledge of
means and variances alone is insufficient, and a description of the full
distribution of molecular counts is necessary. For this we need a (slightly)
more complex model than model 1 that would allow us to incorporate the
non-Poissonian features of constitutive promoters directly into a master
equation formulation.

@@

Our second model of constitutive transcription posits an architecture in which
the promoter is either empty or bound by RNAP ~\cite{Phillips2015a,
Phillips2019}. Here,  as shown in model 2 of Figure~\ref{fig2:constit_cartoons},
transcription initiation results in a state transition from the bound to the
unbound state, reflecting the microscopic reality that an RNAP that has begun to
elongate a transcript is no longer available at the start site to begin another.
As shown in Appendix~\ref{sec:non_bursty}, the Fano factor in this model is
given by
\begin{align}
    \nu = 1 -
        \frac{r\rate{k}{P}{+}}
            {\left(\rate{k}{P}{+} + \rate{k}{P}{-} + r\right)
             \left(\gamma + \rate{k}{P}{+} + \rate{k}{P}{-} + r\right)}.
\end{align}
The problem with this picture is that the Fano factor is always $<1$. To make
 contact with the experimental reality of $\nu>1$ as shown in
 Figure~\ref{fig2:constit_cartoons}(B), clearly some corrections will be needed.
 While this model adds an appealing element of microscopic reality, we are
 forced to reject it as the additional complexity is unable to capture the
 phenomenology of interest. Obviously the promoter state does in fact proceed
 through cycles of RNAP binding, initiating, and elongating, but it seems that
 the super-Poissonian noise in mRNA copy number we want to model must be
 governed by other features of the system.

@@

\marginpar{\it RP: very fun next paragraph - I like the argument, need to think more}
In fact, we suspect \textit{any} model in which transcription proceeds through a
multistep cycle must necessarily have $\nu<1$. The intuitive argument compares
the waiting time distribution to traverse the cycle with the waiting time for a
Poisson promoter (model 1) with the same mean time. The latter is simply an
exponential distribution. The former is a convolution of multiple exponentials,
and intuitively should be more peaked with a smaller fractional width than an
exponential with the same mean.
\footnote{This can be made more precise with a result from~\cite{Stewart2007},
who showed that the convolution of multiple gamma distributions (of which the
exponential distribution is a special case) is, to a very good approximation,
also gamma distributed. Using their Eq.~2 for the distribution of the
convolution, with shape parameters set to 1 to give exponential distributions,
the total waiting time distribution has a ratio of variance to squared mean
$\sigma^2/\mu^2 = \sum_i k_i^2/\left(\sum_i k_i\right)^2$, where the $k_i$ are
the rates of the individual steps. Clearly this is less than 1 and therefore the
total waiting time distribution is narrower than the corresponding exponential.}
A less disperse waiting time distribution means transcription initations are
more uniformly distributed in time relative to a Poisson process. Hence the
distribution of mRNA over a population of cells should be less variable compared
to Poisson, giving $\nu<1$.
\footnote{This last step, while intuitive, can be argued by analogy to photon
statistics where antibunching gives rise to sub-Poissonian noise~\cite{Paul1982,
Zou1990}. Although loopholes exist, we do not expect they apply for our problem.
Nevertheless we refrain from elevating this cycle/sub-Poissonian equivalence to
a ``theorem.'' \mmnote{Antibunching is the ``obvious'' analogy for me but it'd
be totally out of left field for most readers, any suggestions on a more
comprehensible reference??}}
Regardless of the merits of this model in describing the noise properties of
constitutive transcription initiation, it ultimately fails the simplest
quantitative feature of the data, namely that the Fano factor $> 1$ and hence
we must discard this mechanistic picture and search elsewhere.

@@

While this seems promising, there is a major drawback for our purposes which was
already uncovered by the authors of~\cite{Razo-Mejia2020}: the statistical
inference problem is nonidentifiable, in the sense described in Section 4.3
of~\cite{Gelman2013}. What this means is that it is impossible to infer the
parameters $r$ and $k^-$ from the single-cell mRNA counts data
of~\cite{Jones2014} (as shown in Fig.~S2 of~\cite{Razo-Mejia2020}). Rather, only
the ratio $r/k^-$ could be inferred. In that work, the problem was worked around
with an informative prior on the ratio $k^-/k^+$. That approach is unlikely to
work here, as, recall, our entire goal in modeling constitutive expression is to
use it as the basis for a yet more complicated model, when we add on repression.
But adding more complexity to a model that is already poorly identified is a
fool's errand, so we will explore one more potential model.

@@

The final model we consider is 
inspired by the failure mode of model 4. The key observation above
was that, as found in~\cite{Razo-Mejia2020}, only two parameters, $k^+$ and the
ratio $r/k^-$, could be directly inferred from the single-cell mRNA data
from~\cite{Jones2014}. So let us take this seriously and imagine a model where
these are the only two model parameters. What would this model look like?

@@

Let us now verify this intuition with a precise derivation to check that $r/k^-$
is in fact the mean burst size and to obtain the full burst size distribution.
Consider first a constant, known dwell time $T$ in the active state. Transcripts
are produced at a rate $r$ per unit time, so the number of transcripts $n$
produced during $T$ is Poisson distributed with mean $rT$, i.e.,
\begin{equation}
    P(n\mid T) = \frac{(rT)^n}{n!} \exp(-rT).
\end{equation}
Since the dwell time $T$ is unobservable, we actually want $P(n)$, the dwell
time distribution with no conditioning on $T$. Basic rules of probability theory
tell us we can write $P(n)$ in terms of $P(n\mid T)$ as
\begin{equation}
    P(n) =\int_0^\infty P(n\mid T) P(T) dT.
\end{equation}
But we know the dwell time distribution $P(T)$, which is exponentially
distributed according to
\begin{equation}
    P(T) = k^- \exp(-T k^-),
\end{equation}
so $P(n)$ can be written as
\begin{equation}
    P(n) = k^- \frac{r^n}{n!}
            \int_0^\infty T^n\exp[-(r + k^-)T]\,dT.
\end{equation}
A standard integral table shows $\int_0^\infty x^n e^{-ax}\,dx = n!/a^{n+1}$, so
\begin{equation}
    P(n) = \frac{k^- r^n}{(k^- + r)^{n+1}}
        = \frac{k^-}{k^- + r}
            \left(\frac{r}{k^- + r}\right)^n
        = \frac{k^-}{k^- + r}
            \left(1 - \frac{k^-}{k^- + r}\right)^n,
\end{equation}
which is exactly the geometric distribution with standard parameter
$\theta\equiv k^-/(k^- + r)$ and domain $n \in \{0, 1, 2, \dots\}$ (this is one
of two common conventions for the geometric distribution). The mean of the
geometric distribution, with this convention, is
\begin{align}
\langle n\rangle = \frac{1 - \theta}{\theta}
        = \left(1 - \frac{k^-}{k^- + r}\right)
                    \frac{k^- + r}{k^-}
        = \frac{r}{k^-},
\end{align}
exactly as we guessed intuitively above.

So in taking the limit $r,k^-\rightarrow\infty$, $r/k^-\equiv b$, we obtain a
model which effectively has only a single promoter state, which initiates bursts
at rate $k^+$ (transitions to the active state, in the model 4 picture). The
master equation for mRNA copy number $m$ as derived in
Appendix~\ref{sec:non_bursty} takes the form
\begin{align}
\begin{split}
{d \over dt}p(m,t) = & (m+1)\gamma p(m+1,t) - m\gamma p(m,t) \\
        &+ \sum_{m'=0}^{m-1} k_i p(m',t) \text{Geom}(m-m';b)
         - \sum_{m'=m+1}^\infty k_i p(m,t) \text{Geom}(m'-m;b),
\end{split}
\end{align}
where we use $k_i$ to denote the burst initiation rate, $\text{Geom}(n;b)$ is the
geometric distribution with mean~$b$, i.e., $\text{Geom}(n;b) =
\frac{1}{1+b}\left(\frac{b}{1+b}\right)^n$ (with domain over nonnegative
integers as above). The first two terms are the usual mRNA degradation terms.
The third term enumerates all ways the system can produce a burst of transcripts
and arrive at copy number $m$, given that it had copy number $m'$ before the
burst. The fourth term allows the system to start with copy number $m$, then
produce a burst and end with copy number $m'$. In fact this last sum has trivial
$m'$ dependence and simply enforces normalization of the geometric distribution.
Carrying it out we have
\begin{equation}
\begin{split}
{d \over dt} p(m,t) = & (m+1)\gamma p(m+1,t) - m\gamma p(m,t) \\
        &+ \sum_{m'=0}^{m-1} k_i p(m',t) \text{Geom}(m-m';b)
            - k_i p(m,t),
\end{split}
\end{equation}
We direct readers again to Appendix~\ref{sec:gen_fcn_appdx} for further details.
This improves on model 4 in that now the parameters are easily inferred, as we
will see later, and have clean interpretations. The non-Poissonian features are
attributed to the empirically well-established phenomenological picture of
bursty transcription.

@@

Consider the time delay between the first and final RNAPs in a burst initiating
transcription (\textit{not} the time to complete transcripts, which potentially
could be much longer.) If this timescale is short compared to the typical search
timescale of transcription factors, then all is well. The estimates from
deHaseth et.\ al.~\cite{DeHaseth1998} put RNAP's diffusion-limited on rate
around $\sim\text{few}\times10^{-2}~\text{nM}^{-1}~\text{s}^{-1}$ and polymerase
loading as high as $1~\text{s}^{-1}$. Then for reasonable burst sizes of $<10$,
it is reasonable to guess that bursts might finish initiating on a timescale of
tens of seconds or less (with another 30-60 sec to finish elongation, but that
does not matter here). A transcription factor with typical copy number of order
10 (or less) would have a diffusion-limited association rate of order
$(10~\text{sec})^{-1}$ \cite{Hammar2014}. Higher copy number TFs tend to have
many binding sites over the genome, which should serve to pull them out of
circulation and keep their effective association rates from rising too large.
Therefore, there is \textit{perhaps} a timescale separation possible between
transcription factor association rates and burst durations, but this assumption
could very well break down, so we will have to keep it in mind when we infer
repressor rates from the Jones et.\ al.\ single-cell mRNA counts data
later~\cite{Jones2014}.
